---
title: "The Leslie Matrix - revisited"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

A population of fish we're interested in has four life stages: eggs (E), fry (F), juvenile (J), breeding adult (A). You are told that annually:

-   Each breeding adult will produce, on average, 600 eggs
-   5% of eggs will survive to become fry
-   20% of fry will survive to become juveniles
-   16% of juveniles will survive to become adults
-   Adult survival rate year to year is 90%

Recall the Leslie matrix from EDS 212. Matrix multiplication (using the dot product `%*%` of the matrix with the populations at each life stage) yields the population structure at the following year. - \# changes inpopualtions based on previous lifestages \### Create the Leslie matrix:

```{r}
# Make the Leslie Matrix:
fish_leslie <- matrix(c(0, 0, 0, 600, 0.05, 0, 0, 0, 0, 0.20, 0, 0, 0, 0, 0.16, 0.9), nrow = 4, ncol = 4, byrow = TRUE)
 
# Check it out: 
fish_leslie
```

-   no hold over from previous year so one data point in each row The initial population structure is given by:

```{r}
# Initial population structure (Roe, Fry, Juvenile, Adult), initial condition, dot product between amtrix system of equations and give values for next populations, dot product of matrix from pop structure from cycle 1 for cycle 2 and sooo on (this is an iterative process!)
fish_year0 <- c(5000, 8000, 600, 400)
```

Add line-by-line comments wheres prompted in the code below, explaining what each is doing. Make sure that you are talking with your neighbors about what is really being created at each step.

```{r}
# Model projections

# ------ WHAT HAPPENS HERE? ------ #
# ------ comment line-by-line ---- #

# The line below... each year is a cycle, creating this sequence,rather than saying time 1 then time 2 just creates a sequence

time_yr <- seq(from = 0, to = 8, by = 1)

# The line below... year projections creates empty matrix with dimensions of columns with initial condition 4 and rows are length time years, have 9 values (9x4 matrix)
proj_year <- matrix(ncol = length(fish_year0), nrow = length(time_yr))

# The line below...have year 0 population structure, 5000,8000 etc and want that in row 1, [row, column], leave any term empty across an term in axis, 1st row of matrix to be popualted by initial condition of fish year 0, look at proj_year! still have no values for rest of matrix, system for linear equations for estimating next year
proj_year[1, ] <- fish_year0

# The for loop below... dont use seq bbecause not wanting to populatae at row 1, starting at row 2, first term is set, up to length time_yr, dot product for system of linear equations, take dot product of matrix, iterate to find pop at next cycle, take dot product of nature of coefficients roe..,dot product of coefficient matrix and row containing vector... pop structure from previous cycle, i start at 2 take dot product of fish leslie matrix of first row (Proj_year[i-1]) gives you proj estimates at year 1 based on year 0 data, placed in row 2 proj_year[i, ], every iteration rew row is popualted at next cycle from dot pproduct from matrix from dot product at previous year... until end t = 8 years 

for (i in 2:length(time_yr)) {
  proj_year[i,] <- fish_leslie %*% proj_year[i-1,]
}

#Datawrangling

# The rest is wrangling & visualization (run to check it out): re assignm column names to these names
colnames(proj_year) <- c("eggs", "fry", "juvenile", "adult")

proj_df <- data.frame(time_yr, proj_year) %>% 
  pivot_longer(cols = -time_yr, names_to = "lifestage", values_to = "stage_population")
#converts formatin dataframe increasing rows and decreasing number of columns, one variabel lifestage over 4 colum creates single column for lifestage and second column 
#take column names melt columns names, nto one, names_to = "species", laues _to = "count

ggplot(data = proj_df, aes(x = time_yr, y = stage_population)) +
  geom_line(aes(color = lifestage)) +
  scale_y_log10()
```

demo \<- data.frame(year = c(1995, 2000, 2005), cats = c(2,4,6), dogs = c(10,14,12))

Test different combinations of survival / reproduction and starting population structures to see what happens!

ggplot(data = )

#string to title convert string to any case you want

food = 10, call food with no questes, just saying the name

vs

this string("food")
